---
title: "Marvel vs Superman"
author: "Yue"
date: "8/18/2017"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(compare)
library(lubridate)
library(rvest)
library(stringr)
```

```{r}
url_data_marvel <- 
  "https://en.wikipedia.org/wiki/List_of_films_based_on_Marvel_Comics"
url_data_dc <- 
  "https://en.wikipedia.org/wiki/List_of_films_based_on_DC_Comics"

```

First, let's get the Marvel data. I'm pulling this from Wikipedia. We'll do something
similar with the DC data.

```{r, WARNING=FALSE}
marvel <-
  read_html(url_data_marvel) %>% 
  html_node(css = "#mw-content-text > div > table:nth-child(28)") %>% 
  html_table(fill = TRUE)

colnames(marvel)[5:8] <- str_c(colnames(marvel)[5:8], " ", marvel[1,5:8])
marvel <-
  marvel %>% 
  slice(-1) %>% 
  select(
    title = Title,
    distributor = `Distributor(s)`,
    release = `Release date\n(United States)`,
    budget = `Budget\n(in millions)`,
    box_opening = `Box office gross[28] Opening weekend\n(North America)`,
    box_na = `Box office gross[28] North America`,
    box_world = `Box office gross[28] Worldwide`
  ) 

marvel$release <- marvel$release %>% parse_date("%B %d, %Y")
marvel$box_opening <- parse_number(marvel$box_opening)
marvel$box_na <- parse_number(marvel$box_na)
marvel$budget <- parse_number(marvel$budget) * 1e6
marvel$box_world <- parse_number(marvel$box_world)

marvel <-
  marvel %>% 
  filter(!is.na(release))
```

Now, let's get the rating, which is in a separate table. One of the things we
want to do is get rid of some of the ratings with less than 20. It's not significant
enough to say that these reviews mean anything when only 20 people have reviewed it.

```{r, warning=FALSE}
marvel_rating <-
  read_html(url_data_marvel) %>% 
  html_node(css = "#mw-content-text > div > table:nth-child(30)") %>% 
  html_table(fill = TRUE) %>% 
  select(Film, rating = `Rotten Tomatoes`) %>% 
  mutate(
    review_num = str_extract(rating, "\\d+( reviews)"),
    review_num = parse_number(review_num),
    rating = parse_number(rating)
  ) %>% 
  filter(review_num > 20)
```

Now we'll join those two tables together. Here's where things get tricky: we have
a few movies in the marvel universe that are repeats, and some of these repeats
have information on the box office, and some don't. Because the ratings table
doesn't have a date next to it, there are two Fantastic Four's that are labeled with 
(2005) and (2015). And what's worse, we have two The Punisher, one of them without
any box office information. Joining these two tables will be annoying... 
So first, I'm going to make sure we don't miss The Fantastic Four by counting all 
of the Fantastic Four's in the first table. Then I'm going to see which one is NA
(since all the movies have ratings, but not necessarily box office info) and 
add the year. This should take care of everything!

```{r}
marvel <-
  marvel %>% 
  count(title) %>% 
  left_join(marvel, ., by = "title") %>% 
  mutate(
    title = ifelse(n > 1, str_c(title, " (", lubridate::year(marvel$release), ")"), title),
    universe = "marvel"
  ) %>% 
  left_join(marvel_rating, by = c("title" = "Film")) %>% 
  mutate(title = ifelse(is.na(rating), str_c(title, " (", lubridate::year(marvel$release), ")"), title)) %>% 
  select(-rating, -n, -review_num) %>% 
  left_join(marvel_rating, by = c("title" = "Film"))
```

Now, let's get the DC data. This is a lot less tricky than the Marvel movies, phew!

```{r, warning=FALSE}
dc <-
  read_html(url_data_dc) %>% 
  html_node(css = "#mw-content-text > div > table:nth-child(33)") %>% 
  html_table(fill = TRUE)
colnames(dc)[5:7] <- str_c(colnames(dc)[5:7], " ", dc[1,5:7])

dc <- 
  dc %>% 
  slice(-1) %>% 
  select(
    title = Title,
    distributor = `Distributor(s)`,
    release = `Release date\n(United States)`,
    budget = `Budget\n(millions)`,
    box_opening = `Box office gross[8] Opening weekend\n(North America)`,
    box_na = `Box office gross[8] North America`,
    box_world = `Box office gross[8] Worldwide`
  ) 

dc$release <- dc$release %>% parse_date("%B %d, %Y")
dc$box_opening <- parse_number(dc$box_opening)
dc$box_na <- parse_number(dc$box_na)
dc$budget <- parse_number(dc$budget) * 1e6
dc$box_world <- parse_number(dc$box_world)

dc <-
  dc %>% 
  filter(title != "Total")
```

And now, the ratings!

```{r}
dc_rating <-
  read_html(url_data_dc) %>% 
  html_node(css = "#mw-content-text > div > table:nth-child(35)") %>% 
  html_table(fill = TRUE) %>% 
  select(Film, rating = `Rotten Tomatoes`) %>% 
  mutate(
    review_num = str_extract(rating, "\\d+( reviews)"),
    review_num = parse_number(review_num),
    rating = parse_number(rating)
  ) %>% 
  filter(review_num > 20)
```
And join the two (really easily!)

```{r}
dc <-
  dc %>% 
  left_join(dc_rating, by = c("title" = "Film")) %>% 
  mutate(universe = "dc") 
```

And now let's put that into one large table!
```{r}
all_universe <-
  dc %>% 
  merge(marvel, all = TRUE) %>% 
  mutate(rating = rating / 100)
```

Let's first do a quick sketch of some of our key variables. What's the distribution
of the ratings?

```{r, fig.height=10}
all_universe %>% 
  filter(!is.na(rating)) %>% 
  ggplot(aes(forcats::fct_reorder(title, rating), rating)) +
  facet_grid(universe ~ ., scale = "free_y", space = "free") +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = scales::percent)
```

We can see that both of the universes have a generally a wide variety of good ratings.
The distributions almost look similar. Marvel definitely seems to have more positive
ratings, but that's also skewed by the fact that Marvel has a larger selection of
movies.

How does it do in box office?

```{r, fig.height=10}
all_universe %>% 
  filter(!is.na(box_world)) %>% 
  ggplot(aes(forcats::fct_reorder(title, box_world), box_world)) +
  facet_grid(universe ~ ., scale = "free_y", space = "free") +
  geom_col() +
  coord_flip() 
```

Well, it seems like Marvel is the top dog in terms of box office, especially 
with its most popular movies, but there's
a good amount of lesser known movies too. Like, what's Howard the Duck?

Also, how do these movies fare worldwide vs in the US. What makes more money outside
the US comparatively.

```{r, fig.height=10}
all_universe %>% 
  mutate(world_us_ratio = box_world / box_na) %>% 
  ggplot(aes(forcats::fct_reorder(title, world_us_ratio), world_us_ratio)) +
  facet_grid(universe ~ ., scale = "free_y", space = "free") +
  geom_col() +
  coord_flip() 
```

Well, it definitely seems like some of the DC movies first of all, didn't get
(any) traction abroad. It also seems like the Marvel universe does really well outside
the US in general. Only a few movies, like Constantine, Batman v Superman, Dark Knight,
etc. has had popularity abroad. It seems like most Marvel movies made at least half
its profit outside the US.

And also, I'm curious whether ratings help boost the movies too.

```{r}
all_universe %>%  
  ggplot(aes(rating, box_world)) +
  geom_smooth(se = FALSE, color = "black") +
  geom_point(aes(color = universe)) +
  scale_x_continuous(labels = scales::percent) +
  ggrepel::geom_text_repel(
    data = all_universe %>% group_by(universe) %>% filter(box_world == max(box_world)), 
    aes(label = title)
  ) +
  ggrepel::geom_text_repel(
    data = all_universe %>% filter(!is.na(box_world) & !is.na(rating)) %>% group_by(universe) %>% filter(box_world == min(box_world)), 
    aes(label = title)
  ) +
  labs(y = "Box office gross worldwide ($)") +
  scale_color_manual(name = "universe", values = c("blue", "red"))

```

Well, it's definitely quite a mix, but it's pretty easy to see that there's a moderate
correlation with good ratings. Not sure if it's causation, but I guess a good
word of mouth will probably help. It seems like a good rating might help out a movie,
but it doesn't guarantee it. But it's pretty clear that some of the bad-rated movies
also did pretty bad in movie theaters. 

What about the relationship between budget and box office?

```{r}
all_universe %>% 
  ggplot(aes(budget, box_world)) +
  geom_point(aes(color = universe)) +
  geom_abline() +
  scale_color_manual(name = "universe", values = c("blue", "red")) +
  ggrepel::geom_text_repel(
    data = all_universe %>% filter(!is.na(box_world) & !is.na(budget)) %>% filter(box_world < budget), 
    aes(label = title)
  ) 
```

It seems like a good amount of DC movies do quite terribly. Most of the movies
that are below the black line (aka making less money than it spent) are DC movies.
On the other hand, we're seeing a lot more success with Marvel than DC.

But what about the distributors? We see that DC universe is mostly dominated by
Warner Bros, while the distributors for Marvel movies are far and wide. Especially
for Walt Disney , 20th Century Fox, and Sony Pictures. And it seems like Walt Disney
does the best at making back some of the money that they spent. We can see that
Disney makes back almost four times the amount spent per movie! Sony Pictures also
does pretty well... It seems like Tri-Star and Lionsgate though, haven't done too
well with the universe.

```{r}
all_universe %>% 
  group_by(distributor) %>% 
  filter(!is.na(budget)) %>% 
  summarize(
    n = n(),
    avg_budget = sum(budget) / n,
    avg_box_office = sum(box_world) / n,
    avg_net_gain = avg_box_office - avg_budget,
    rate = avg_net_gain / avg_budget
  ) %>% 
  arrange(desc(rate))
```

But what about seasons and times of release? Let's first get seasons.

```{r}
get_season <- function(date) {
  mmdd <- 100 * month(date) + day(date)
  cuts <- cut(mmdd, breaks = c(0, 319, 620, 921, 1220, 1231))
  levels(cuts) <- c("Winter", "Spring", "Summer", "Fall", "Winter")
  return(cuts)
}

all_universe <-
  all_universe %>% 
  mutate(season = get_season(release))

all_universe %>% 
  ggplot(aes(release, box_world)) +
  geom_point(aes(color = season)) +
  geom_smooth(se = FALSE) +
  scale_color_manual(name = "season", values = c("light blue", "light green", "yellow", "orange"))

all_universe %>% 
  filter(!is.na(budget)) %>% 
  ggplot(aes(release, budget)) +
  geom_point(aes(color = universe)) +
  ggrepel::geom_text_repel(
    data = all_universe %>% filter(!is.na(budget)) %>% group_by(universe) %>% filter(budget == min(budget)),
    aes(label = title)
  ) +
  scale_color_manual(name = "universe", values = c("blue", "red"))

```

We see that most movies are out in spring, then summer and winter. Fall is not quite
a popular season. It also seems like over time, movies are making a lot more money,
but it's not true for every movie. Some of the more recent movies are also flops.
By universe, we can see that Marvel in general is has less "flops" than DC. Of course,
"flops" should also be considered by the amount they spent, but most of the movies
with little box office worldwide is from the DC universe.

```{r}
all_universe <-
  all_universe %>% 
  mutate(profit_ratio = box_world / budget)

all_universe %>% 
  filter(!is.na(profit_ratio)) %>% 
  ggplot(aes(release, profit_ratio)) +
  geom_point(aes(color = universe)) +
  ggrepel::geom_text_repel(
    data = all_universe %>% filter(!is.na(profit_ratio)) %>% group_by(universe) %>% filter(profit_ratio == max(profit_ratio)),
    aes(label = title)
  ) +
  scale_color_manual(name = "universe", values = c("blue", "red")) +
  geom_hline(yintercept = 1) +
  ggrepel::geom_text_repel(
    data = all_universe %>% filter(!is.na(profit_ratio)) %>% group_by(universe) %>% filter(profit_ratio == min(profit_ratio)),
    aes(label = title))
```

OK, let's look at some profit ratios. What makes a lot more money than its budget?
Well, it seems like Batman, which didn't seem like a hit just looking at box office
is the most successful movie in terms of budget returns for DC. And Deadpool also
made way more than its budget. But we can see that Steel (who's even heard of that)
and Punisher just didn't do too well recently.


Now, I'm wondering if there are particular seasons that are more popular for release,
and if seasons make a difference in how these movies do in box office. 

```{r}
all_universe %>% 
  ggplot(aes(season, fill = universe)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(name = "universe", values = c("blue", "red"))

all_universe %>% 
  mutate(month = lubridate::month(release, label = TRUE)) %>% 
  ggplot(aes(month, fill = universe)) +
  geom_bar(position = "dodge") +
  scale_fill_manual(name = "universe", values = c("blue", "red"))
```

Well, it seems like for both universe, Spring is the most popular, followed by
Summer. Fall is just not a popular time for DC/Marvel movies. The month breakdown
is even more interesting: we see that Marvel REALLY likes to release movies in May,
and DC isn't even trying to compete (or something like that.). While June is REALLY
popular for the DC universe. Also, there's been no DC movies released in January, ever.

But of course, those numbers are a bit skewed by the number of movies. Marvel has
significantly more movies released, so I want to look at proportions instead.
We can see that Spring is still the most popular time for both movies. 50% of Marvel
movies come out in Spring! Summer, surprisingly is less popular. I'm defining Summer 
to be after late June, but it seems like July-August period is still not too popular
for Marvel movies.

```{r}
all_universe %>% 
  count(universe, season) %>% 
  group_by(universe) %>% 
  mutate(season_prop = n / sum(n)) %>% 
  ggplot(aes(season, season_prop, fill = universe)) +
  geom_col(position = "dodge") +
  scale_fill_manual(name = "universe", values = c("blue", "red"))
```

```{r}
all_universe %>% 
  mutate(month = lubridate::month(release, label = TRUE)) %>% 
  count(universe, month) %>% 
  group_by(universe) %>% 
  mutate(month_prop = n / sum(n)) %>% 
  ggplot(aes(month, month_prop, fill = universe)) +
  geom_col(position = "dodge") +
  scale_fill_manual(name = "universe", values = c("blue", "red"))
```

But, here we can see that more than 40% of DC movies come out in June! And more than
a third of Marvel movies come out in May. So expect the next Avengers or Captain
America to be in May, and the next batman to be in June...

